---
title: "Data Harmony"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
params:
  echo: no
  rootdir: ~/founder_diet_study
---

```{r}
harmonizeddir <- file.path(params$rootdir, "HarmonizedData")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = harmonizeddir)
```

```{r}
library(dplyr)
library(tidyr)
library(readxl)
library(readr)
library(stringr)
library(foundr) # github/byandell/foundr
```

```{r}
#harmonizeddir <- "."
rawdir <- "../RawData"
getwd()
```

```{r}
for(filename in (filenames <- list.files("R", full.names = TRUE))) {
  source(filename)
}
filenames
```

# Harmonize Data

Data for this repository are identified by `data/RawData/source.csv`,
which is saved with the repo.

Condition name and normalize

```{r}
condition_name <- "diet"
normalize <- FALSE
```
 
```{r}
links <- read.csv(file.path("..", "RawData", "source.csv"), fill = TRUE) %>%
  mutate(address = file.path(rawdir, address))
```

## Annotation

The experiment annotation file relates the `number` (identifier for `animal`) to the `sex` and `diet`. Needed for `LivRna`.

```{r}
annotfile <- linkpath("annot", links)
excel_sheets(annotfile)
annot <- read_excel(annotfile) %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```

# Datasets

## Physio: Physiological traits

```{r}
harmonize("Physio", links, PhysioHarmony, annot, condition_name = "diet",
          sheet = 2,
          rename_function = function(x) {
            rename(x, KetoneBodies = "KB_14wk")
          },
          normalize = normalize)
```

## PlaMet: Plasma metabolites

These are data hand-curated by Mark from data provided by Qiushi
using an Excel script and saved as CSV.

```{r warning = FALSE}
harmonize("PlaMet", links, EnrichHarmony, annot, skiprows = 4,
          condition_name = condition_name, normalize = normalize)
```

Following are old versions not to be used now

```{r warning = FALSE, eval = FALSE}
harmonize("PlaMet0", links, MetHarmony, annot, skiprows = 4,
          condition_name = condition_name, normalize = normalize)
harmonize("PlaMet120", links, MetHarmony, annot, skiprows = 3,
          condition_name = condition_name, normalize = normalize)
harmonize("PlaMetAll", links, MetHarmony, annot, skiprows = 4,
          condition_name = condition_name, normalize = normalize)
```

## LivMet: Liver metabolites

```{r warning = FALSE}
harmonize("LivMet", links, MetHarmony, annot, skiprows = 3,
          condition_name = condition_name, normalize = normalize)
```

## LivRna: Liver mRNA expression

These data have been run through variance stabilizing transform and
do not need further normalization.

```{r warning = FALSE}
harmonize("LivRna", links, LivRnaHarmony, annot,
          condition_name = condition_name, normalize = normalize)
```

## LivIso: Liver mRNA isoform data

```{r}
harmonize("LivIso", links, LivRnaHarmony, annot, isoform = TRUE,
          condition_name = condition_name, normalize = normalize)
```

Special situation: get stats for all data.

```{r eval=FALSE}
LivIsoStats <- harmonize("LivIso", links, LivRnaHarmony, annot, isoform = TRUE,
          condition_name = condition_name, normalize = normalize,
          drop_na_pvalue = FALSE)
```

```{r eval=FALSE}
saveRDS(LivIsoStats, "~/Documents/Research/attie_alan/FounderDietStudy/data/LivIsoStats.rds")
```

## Lipid: Liver Lipids

```{r}
harmonize("Lipid", links, LipidHarmony, annot,
                sheetNum = 1  ,
                skipNum = 0,
                minCol=2,
                maxCol=16,
          condition_name = condition_name, normalize = normalize)
```

Note that `sampleKey` had 2 errors in it, which have been corrected. (B9-9 -> B6-9; 129-5 -> 129-15)

```{r eval=FALSE}
sampleKey <- read_excel("../Primary data from Mark/192 livers sample key.xlsx")
sampleKey[sampleKey[1] == 60, 2] <- "B6-9"
sampleKey[sampleKey[1] == 167, 2] <- "129-15"
```

```{r eval=FALSE}
harmonize("OldLipid", links, OldLipidHarmony, annot,
                sheetNum = 1  ,
                skipNum = 1,
                minCol=2,
                maxCol=5,
                charCols=c("mouse_id", "Total Protein (µg)", "Tissue Mass (mg)", "strain", "animal", "sex", "condition","Name"),
                sampleKey=sampleKey,
          condition_name = condition_name, normalize = normalize)
```

```{r eval = FALSE}
dir(".")
LipidData <- readRDS("Lipid/LipidData.rds")
LipidData <- LipidData %>% mutate(animal = as.character(animal))
saveRDS(LipidData, "Lipid/LipidData.rds")
```

```{r eval = FALSE}
LipidStats <- readRDS("Lipid/LipidStats.rds")
keepTraits <- unique(LipidStats$trait)
LipidSignal <- partition(readRDS("Lipid/LipidData.rds"))
LipidSignal <- 
  dplyr::filter(
      partition(readRDS("Lipid/LipidData.rds")),
      .data$trait %in% keepTraits)
saveRDS(LipidSignal, "Lipid/LipidStats.rds")
```

## Enrich: Plasma enrichment over time

```{r}
harmonize("Enrich", links, EnrichHarmony, annot,
          condition_name = condition_name, normalize = normalize)
```
  
```{r}
harmonize("PlaEnrich", links, Enrich3Harmony, annot,
          condition_name = condition_name, normalize = normalize)
```

```{r eval = FALSE}
dir(".")
PlaEnrichData <- readRDS("PlaEnrich/PlaEnrichData.rds")
PlaEnrichData <- PlaEnrichData %>% mutate(animal = as.character(animal))
saveRDS(PlaEnrichData, "PlaEnrich/PlaEnrichData.rds")
```

```{r}
harmonize("LivEnrich", links, LivEnrichHarmony, annot,
          condition_name = condition_name, normalize = normalize)
```

## Module: WGCNA module eigentraits

See <WGCNA.Rmd> and <WGCNAmixed.Rmd> for construction of `traitModule`.

```{r}
traitModule <- readRDS(file.path("traitModule.rds"))
```

```{r}
harmonize("Module", links, foundr::moduleHarmony,
          traitModule[!grepl("MixMod",names(traitModule))],
          condition_name = condition_name, normalize = normalize)
```

```{r}
harmonize("MixMod", links, foundr::moduleHarmony,
          traitModule["MixMod"],
          condition_name = condition_name, normalize = normalize)
```

# Bind traits into RDS objects for shiny tools

Standard `FounderDietStudy`:

```{r}
bind_traits(c("Physio","PlaMet","LivMet","Lipid","Module","Enrich","PlaEnrich","LivEnrich"))
```


Special `FounderLiverDietStudy`:

```{r}
bind_traits(c("Physio","LivMet","Lipid","LivRna","LivIso","MixMod"), traitRoot = "liver")
```

Small tester

```{r}
bind_traits(c("Physio","Mixmod"), traitRoot = "PhysMix")
```

